### 2020-05-27 2020_EFM_MPM_Eigensoftening.tex
PNA+DM_20_05_27

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx} % figuras
% \usepackage{subfig}
\usepackage{subfigure} % subfiguras
\usepackage{multirow} 
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init macros math

\usepackage[fleqn,tbtags]{amsmath}
\usepackage{mathtools}
\usepackage{amsthm,amssymb}
\usepackage{xstring}

\newcommand{\vect}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}
% End macros math
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \usepackage{algorithm,algcompatible}

\usepackage{blindtext}
\usepackage[linesnumbered,boxed]{algorithm2e}

\SetAlCapSkip{1em}

\SetKwInput{KwInput}{Input}
\SetKwInput{KwOutput}{Output}

% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
% \algnewcommand\INPUT{\item[\textbf{Input:}]}%
% \algnewcommand\OUTPUT{\item[\textbf{Output:}]}%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init glossaries
\usepackage[acronym]{glossaries}

\makeglossaries

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}

\newglossaryentry{domain}
{
  name=$\varOmega$,
  description={Continuum domain}
}

\newglossaryentry{contour}
{
  name=$\partial\varOmega$,
  description={Boundary of the continuum domain $\varOmega$. Also
    defined as $\Gamma$}
}

\newglossaryentry{dirichlet-boundary}
{
  name=$ \Gamma_d$,
  description={Essential or Dirichlet boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Nemann boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{a}
{
  name=$\vect{a}$,
  description={First order tensor which describes the acceleration field}
}

\newglossaryentry{v}
{
  name=$\vect{v}$,
  description={First order tensor which describes the velocity field}
}

\newglossaryentry{u}
{
  name=$\vect{u}$,
  description={First order tensor which describes the displacement field}
}

\newglossaryentry{x}
{
  name=$\vect{x}$,
  description={First order tensor which describes the global coordinates field}
}

\newglossaryentry{xi}
{
  name=$\vect{\xi}$,
  description={First order tensor which describes the local coordinates field}
}

\newglossaryentry{stress}
{
  name=$\tens{\sigma}$,
  description={Second order tensor which means the Cauchy stress tensor}
}

\newglossaryentry{strain}
{
  name=$\tens{\varepsilon}$,
  description={Second order tensor which means the Cauchy strain tensor}
}

\newglossaryentry{Constitutive}
{
  name=$\tens{D}$,
  description={Four order tensor which means the constitutive response
    of the material}
}

\newglossaryentry{LME-beta}
{
  name=$\beta$,
  description={Regularization or thermalization parameter of the
    LME$_{\beta}$ Pareto set}
}

\newglossaryentry{rho}
{
  name=$\rho$,
  description={Describes the scalar density field}
}

\newglossaryentry{fc}
{
  name=f$_c$,
  description={Compressive strenght}
}

\newglossaryentry{ft}
{
  name=f$_t$,
  description={Tensile strenght}
}

\newglossaryentry{Gf}
{
  name=G$_F$,
  description={Specific fracture energy}
}

\newglossaryentry{E}
{
  name=E,
  description={Elastic modulus}
}

\newglossaryentry{nu}
{
  name=$\nu$,
  description={Poisson ratio}
}
\begin{itemize}
\item 
\end{itemize}

\newglossaryentry{gamma-npc}
{
  name=$\gamma$,
  description={Time integration parameter for the \acrshort{npc} algorithm}
}

\newglossaryentry{gamma-lme}
{
  name=$\widehat{\gamma}$,
  description={Dimensionless parameter to control the value of the
    thermalization parameter $\beta$ of the \acrshort{lme} shape functions.}
}

% End glossaries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\etal}{
  \textit{et al.}
}

\newcommand{\PNA}[1]{
  \textcolor{red}{{#1}}
}

\newcommand{\MMP}[1]{
  \textcolor{blue}{{#1}}
}

\newcommand{\DM}[1]{
  \textcolor{green}{{#1}}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \journal{Engineering Fracture Mechanics}
\journal{Computers and Structures}

\modulolinenumbers[5]

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{LME-MPM applied to quasi-brittle fracture.}

% \title{An enhanced EigenMPM }

%% Group authors per affiliation:
\author{
Miguel Molinos$^a$\footnote{Corresponding author: m.molinos@alumnos.upm.es},
Pedro Navas$^a$\footnote{Corresponding author: p.navas@upm.es},
Diego Manzanal$^a$,
and Manuel Pastor$^a$
 }
 \address{
 $^a$ ETSI Caminos, Canales y Puertos, Universidad Polit\'ectnica de Madrid.\\
 c. Prof. Aranguren 3, 28040 Madrid, Spain
}

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for the
  \acrshort{mpm}. The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the well performance of the method when challenging
  applications are to modeled.
\end{abstract}

\begin{keyword}
Quasi brittle fracture \sep Local-\textit{Max-Ent} approximation \sep
Material Point Method \sep Solid Dynamics
\end{keyword}

\end{frontmatter}

\linenumbers

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The presence of cracks is a violation of the continuity requirement of the
\acrfull{fem}. To overcome this issue, numerous numerical
artifacts have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, \PNA{to the gradual insertion of}
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries, or handling arbitrary crack paths by level set
representation of the fracture surface~\cite{Belytschko_03}. The
simulation of fracture propagation in a more accurate and effective
way can be considered as one of the \PNA{ongoing goals} of the development of novel spatial discretization methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
\acrfull{efgm} \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013}, the
\acrfull{sph} \cite{Wang_2020,Wang_2019}, the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others. \\

Regarding \acrshort{mpm}, fracture \PNA{has been treated numerically through two different ways}. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003,Nairn_2006}. \PNA{The basis of this methodology consist of the removal of} the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. 
On the other hand, the \PNA{foundation} of the second approach is the introduction of failed
material points to describe the crack evolution. In these methods, the
formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} and
Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}.\\

Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
crack sets are approximated by means of eigen-deformations, which
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle facture
was carried out by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. \PNA{In this technique, the concept of ``erosion'' of
the material point is depicted, being the material point able to remain intact or become completely failed or eroded, when no loading
capacity is bore}. This method has been successfully applied to simulate high
complex phenomena such as dynamic fragmentation of metals
\cite{Li_2015}. Recently, Zhang {\it et al.}
\cite{Zhang_EE_2020} adopted the eigenerosion to resolve the dynamic
fracture of brittle materials in the \acrshort{mpm}
framework. Nonetheless, when quasi-brittle materials are simulated
though this approach, the results exhibit an overestimation of tensile
stress and strain peaks. Furthermore, since the work of Zhang {\it et al.}
\cite{Zhang_EE_2020} is made with conventional \acrshort{mpm}/\acrshort{gimp}, stress oscillations are to be faced.\\

To overcome the limitations observed in the EigenMPM
\cite{Zhang_EE_2020}, the present research proposes the eigensoftening
algorithm developed by Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} for the \acrshort{otm} framework and
engineered for quasi-brittle materials. Inspired in the concept of the
crack band model \cite{Bazant83}, since energy dissipation is performed through the softened (or failed) volume, this methodology is able to capture the gradual rather than abrupt
dissipation of the fracture energy. \PNA{Moreover, in order to mitigate stress
oscillations as well as to deal with tensile instabilities, such those that
occurs when particles crosses an element boundary, the work of Molinos~\etal~\cite{Molinos2020} is followed. The traditional \acrshort{mpm} technique is enhanced through the \acrfull{lme} approximation technique
\cite{Arroyo2006} for the spatial discretization as well as a explicitly \acrshort{mpm} \acrfull{npc} scheme is derived for the time discretization.}\\

The paper is structured as follows. First meshfree methodology,
eigenerosion and eigensoftening algorithms are presented in Section
\ref{sec:2}. Then, both approaches are compared and verified by means of
comparisons with analytical and experimental results in Section
\ref{sec:3}. Finally, relevant conclusions are exposed in Section \ref{sec:4}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}

The aim of this section is to describe and introduce some special
techniques required to face the fracture problem under the \acrshort{mpm}
framework. In consequence, this section is structured as follows: first, in
\ref{sec:2.1}, the \acrfull{npc} algorithm for the \acrshort{mpm}
will be exposed; next the \acrshort{lme} approximants are
introduced in \ref{sec:2.2} as an accurate alternative technique to
interpolate data between particles and nodes, and finally fracture
algorithms based on the eigendeformation concept are presented in
\ref{sec:2.3}.

\subsection{The \acrshort{mpm} time integration : A \acrlong{npc}  scheme}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian
method where moving material points, henceforth in this research,
particles ($\square_p$) will carry on all the physical information of the local state ($\sigma_p,
\varepsilon_p$) and a set of fixed background nodes ($\square_I$) will
be introduced to compute the balance of momentum equation. Since the
\acrshort{mpm} possesses the advantages of both Lagrangian and
Eulerian descriptions, no element distortion takes place in the
\acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
an explicit predictor-corrector time integration scheme is adopted. The
purpose of this choice is motivated due to its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor $ \vec{\tilde{v}}_I^{k+1}$ is computed through
nodal velocities $\vec{v}_I$ and accelerations $\vec{a}_I$ and evaluated
in the $k$ time step,
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{\tens{m}_I^{k+1}}\ .
\end{equation}
Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and $N_{Ip}^{k}$ is
the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} =
          \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

\subsection{Spatial interpolation technique : \acrlong{lme} approximants}
\label{sec:2.2}
\acrfull{lme} approximation scheme was introduced by Arroyo \& Ortiz
(2006)\cite{Arroyo2006} as a bridge between finite elements and
meshfree methods. The key idea under this interpolation technique is
the interpretation of each nodal value $N_I$ as a
probability. Related with this definition, two important limits are
introduced. First the maximum-entropy (\textit{max-ent}) limit, which
ensures a \textit{unbiased statistical inference} based on the nodal
data as states the Jayne's\cite{Jaynes1957} principle of
\textit{maximum entropy}. And second the Delaunay triangulation that
warranties the \textit{least width} shape function support. To reach
\DM{delete to} a compromise between two competing objectives, a Pareto set is
defined as,
\begin{align*}
  \label{eq:LME-scheme-pareto-set}
  \text{(LME)}_{\beta} \hspace{0.15cm} &\text{For fixed} \hspace{0.15cm}
                                         \vec{x} \hspace{0.15cm} \text{minimise} \hspace{0.15cm} f_{\beta}(\vec{x}_p, N_I) \coloneqq \beta U(\vec{x}_p,N_I) - H(N_I) \\
                                       &\text{subject to}\
                                         \begin{cases}
                                           N_I \ge 0, \hspace{0.15cm} \text{I=1, ..., n} \\[1em]   
                                           \sum\limits_{I=1}^{N_n}{N_I} = 1 \\[1em]   
                                           \sum\limits_{I=1}^{N_n}{N_I \vec{x}_I} = \vec{x} \\
                                         \end{cases}
\end{align*}            
where $H(N_I)$ is the entropy of the system of nodes following the
definition given by Shannon (1948) \cite{Shannon1948}, the shape
function width is defined as $U(\vec{x}_p,N_I) \coloneqq
\sum_I N_I |\vec{x}_p - \vec{x}_I |^2$, and $\beta$ is a
regularization o \textit{thermalization} parameter such that for
Pareto optimal solutions $\beta \in
(0,\infty)$. Notice that $\beta$ has units of $[L]^{-2}$,
therefore it can be controlled by adjusting a dimensionless
parameter\footnote{To avoid confusion with the \gls{gamma-npc} parameter of
  the \acrshort{npc}, the dimensionless parameter defined in Arroyo \&
  Ortiz \cite{Arroyo2006} as $\gamma$ will be represent by
  \gls{gamma-lme} to preserve as much as possible the original notation.},
$\widehat{\gamma}=\beta h^2$, where $h$ is defined as a
suitable measure of the nodal spacing. With the restrictions of the
Pareto set, the unique solution of the local
\textit{max-ent} problem \acrshort{lme}$_\beta$ is,
 \begin{equation}
  \label{eq:LME-p}
N_I^*(\vec{x})=\frac{\exp\left[ -\beta \; |\vec{x}-\vec{x}_I|^2 +
    \vec{\lambda}^* \cdot (\vec{x}-\vec{x}_I) \right] } {Z(\vec{x},\vec{\lambda}^*)},
\end{equation}
where $Z(\vec{x},\vec{\lambda}^*)$ is the \textit{partition function} defined as,
\begin{equation}
  \label{eq:LME-Z}
Z(\vec{x}, {\vec{\lambda}}) = \sum_{I=1}^{N_n}{ \exp \left[ -\beta \; |\vec{x}-\vec{x}_I|^2 + \vec{\lambda} \cdot (\vec{x}-\vec{x}_I)  \right]},
\end{equation}
and a Lagrange multiplier $\vec{\lambda}^*$ such minimise the function $\log
Z(\vec{x}, \vec{\lambda})$.
The traditional way to compute the optimal value of $\vec{\lambda}^*$
is through a Newton-Raphson procedure, or in more challenging
scenarios by a combination of the Newton-Raphson and the Nelder-Mead
Simplex algorithms \cite{Navas2018a}. Nonetheless, since finite
strains are not involved in the present research, Newton-Raphson is an
enough efficient method. For an uniform nodal spacing, $\beta$ can be
considered constant, thus first derivatives of the interpolation technique $\nabla
N^*_I$ can be obtained 
by evaluating the following expression
\begin{eqnarray}
  \label{eq:LME-gradp} 
\nabla N^*_I &=& -N^*_I \,  (\tens{J}^*)^{-1} \,  (\vec{x} - \vec{x}_I)
\end{eqnarray}
where $\tens{J}$ is the Hessian matrix, defined by
\begin{eqnarray}
  \label{eq:LME-J} 
  \tens{J}(\vec{x}, \vec{\lambda},\beta) &\equiv& \frac{\partial
                                                  \vec{r}}{\partial \vec{\lambda}},\\
  \label{eq:LME-r}
  \vec{r}(\vec{x},\vec{\lambda},\beta) &\equiv& \frac{\partial \log{ Z(   \vec{x},\vec{\lambda}})}{\partial \vec{\lambda}}  = \sum_I^{N_n} p_I(\vec{x},\vec{\lambda},\beta) \, (\vec{x} - \vec{x}_I).
\end{eqnarray}
An additional remark concerning the support of the interpolation
function is that, in practice, the value of $N_I$ decay exponentially following $\exp(-\beta \vec{r}
)$. In this sense,
a good practice is to truncate it under a tolerance  (typically around 10$^{-6}$) that would ensure a reasonable
range of neighbours, see \cite{Arroyo2006} for details. This tolerance
defines a limit value of the influence radius to find the neighbour
nodes of a given integration point.

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. \DM{ XXXX I would modify the following  sentence as "The evaluation of both approaches \cite{Navas_2017_ES} show that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately XXXXX}. A
comparison between both in \cite{Navas_2017_ES} shows that the
eigenerosion algorithm significantly overestimates the tensile  
stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. On the other hand, eigensoftening algorithm agrees
very well with experimental results in all the aspects \DM{which aspect?} . Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

The key idea behind the eigenerosion algorithm is the computation of the
energy-release rate attendant to the failure of material point $p$,
\begin{align}
  \label{eq:energy-release-EE}
&G_p^{k+1} = \frac{C_{\epsilon}}{m_p^{k+1}}  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q W_q^{k+1}\\
  \label{eq:mass-EE}
&m_p^{k+1} =  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q  
\end{align}
where $B_{\epsilon}(x_p^{k+1})$ is a $^n$dimensional sphere of radius $\epsilon$
centered at $x_p^{k+1}$. The particles which lie under this sphere are
known as the $\epsilon$-neighborhood of the material point (see
\cite{Pandolfi_2012}). This concept is conveniently sketched in
Figure \ref{fig:Failed-particles}.
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure-Particle-failed}
  \caption{Scheme of a linear cohesive law, where the shaded area is
    $G_f$, $f_t$ is the tensile strength, and $w_c$ is the critical
    opening displacement.}
  \label{fig:Failed-particles}
\end{figure}
%%%%%%%%%%%
Other parameters are the mass of the neighborhood $m_p^{k+1}$, the
current free-energy density per unit mass  $W_q^{k+1}$ and the
normalizing constant $C_{\epsilon}$, \PNA{which also defines the $\epsilon$-neighborhood size, being $\epsilon = C_\epsilon h$.}

The failure criterion consist of considering that a material point is failed
when $G_p^{k+1}$ surpasses the critical energy release rate that
measures the material-specific energy, $G_F$. The convergence of this
approach has been analyzed by Schmidt {\it et al.}
(2009)\cite{Schmidt_2009}, who proved that $G_F$ converges to the Griffith
fracture when the discretization size tends to zero. It is necessary to
point out that, when a material point overpass the critical energy, its
contribution to the internal forces vector is set to zero, but its
contribution to the mass matrix is preserved.\\

As can be noticed, the eigenerosion algorithm relies over an energetic
failure criterion. Because of this, unrealistic stress
concentration (higher than tensile strength) appears in quasi-brittle
materials \cite{Navas_2017_ES}. To overcome this limitation, the
aforementioned authors proposed the concept of eigensoftening to take
into account the gradual failure in quasi-brittle materials. The idea
behind this concept is inspired in the cohesive fracture. This gradual
failure criterion is plotted in figure \label{fig:Damage-ft-wc}, where
a linear decreasing cohesive law is presented to illustrate the
concept earlier described. In the picture, the shaded region represents
the static fracture energy per unit of area, $G_F$. Notice how a
cohesive crack appears when the maximum tensile strength, $f_t$, is
reached. Once the crack, $w_n$, reaches the value of the
critical crack opening, $w_c$, a stress-free crack is
attained. \DM{ see the following sentence: comma missing? Clarified--->} For intermediate values a damage value between zero and one represents the extension to which the material has failed. \DM{ hasta acá }
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure-Damage}
  \caption{Scheme of a linear cohesive law, where the shaded area is
    $G_f$, $f_t$ is the tensile strength, and $w_c$ is the critical
    opening displacement.}
  \label{fig:Damage-ft-wc}
\end{figure}
%%%%%%%%%%%
For the eigensoftening algorithm, a strength criterion for crack
initialization was adopted. Particularly the maximum principal stress
theory for brittle fracture was considered  \DM{ delete by the authors}  by authors
\cite{Navas_2017_ES}. With that purpose \DM{ which purpose?}, the variation of the averaged
strain energy density in the $\epsilon$-neighborhood of the material point
$\vec{x}_p^{k+1}$ can be obtained as
\begin{equation}
  \label{eq:variation-averaged-strain-energy-density}
  \delta W_{\epsilon,p} = \frac{\partial G_p}{C_{\epsilon}} =
  \frac{1}{m_p} \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q \tens{\sigma}_{q,I} \delta \tens{\varepsilon}_q,
\end{equation}
where $\tens{\sigma}_{q,I}$ is the maximum principal stress of each
material point in the $\epsilon$-neighborhood. At this point, an \textit{effective} strain $\varepsilon_q$ is introduced in order to obtain the variation of
the local strain energy as $\delta W_q = \sigma_{q,1}
\delta\varepsilon_q$. Now, with the assumption that the effective
strain of each material point at every time step is constant in the
neighborhood of $\vec{x}_p^{k+1}$, the equation 
\eqref{eq:variation-averaged-strain-energy-density} can be simplified
 to
\begin{equation}
  \label{eq:variation-averaged-strain-energy-density-simpli}
  \delta W_{\epsilon,p} =
  \frac{\delta \tens{\epsilon}_p}{m_p} \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q \tens{\sigma}_{q,I}. 
\end{equation}
Consequently, it is possible to define an equivalent critical stress at the
material point $\vec{x}_p^{k+1}$ as
\begin{equation}
  \label{eq:equivalent-critical-stress}
  \tens{\sigma}_{\epsilon,p} =
  \frac{1}{m_p} \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q \tens{\sigma}_{q,I}, 
\end{equation}
where $m_p$ is the total mass of the $\epsilon$-neighborhood, defined similarly  \DM{ Pondría:  defined similarly TO  the eigenerosion method} than in the eigenerosion methodology (Eq.~\eqref{eq:mass-EE}).
%\begin{equation}
%  \label{eq:averaged-mass}
%  m_p = \sum_{x_q^{k+1} \in B_{\epsilon}(x_p^{k+1})} m_q.
%\end{equation}
This definition \DM{ dos veces definition en la misma oración, podes poner directamente: The equivalent critical stress leads to a definition of the averaged strain energy ... } of the equivalent critical stress leads to a
definition of the averaged strain energy in terms of the averaged
strain as $\delta W_{\epsilon,p} =
 \tens{\sigma}_{\epsilon,p}\ \delta\varepsilon_p$. The softening behaviour is
activated once $\tens{\sigma}_{\epsilon,p}^{k+1}$ surpasses the
tensile strength, $f_t$. This consists of a reduction of the internal
forces as, 
 \begin{equation}
   \label{eq:f-int-damaged}
   f^{int}_I = \sum_p (1 - \chi_p)\ \tens{\sigma}_{p}^{k+1} \cdot
   \Grad{N_{Ip}}\ \Omega_p
 \end{equation}
where $\chi_p$ and $\Omega_p$ are respectively the damage or softening
variable and the volume for each material point $p$. \PNA{$\chi_p$ is any function, taking
values between zero (an intact material) and one (completely failed
material points), that relates the crack opening and the residual strength. To calculate it, typical shapes of cohesive laws such as linear, bilinear or exponential, may be employed.} For the case of a linear softening such the sketched one in the Figure \ref{fig:Damage-ft-wc}, $\chi_p$ is computed as,
 \begin{equation}
   \label{eq:damaged-variable-chi}
   1 - \chi = \frac{f_n}{f_t} = 1 - \frac{w_n}{w_c}\ \rightarrow\ \chi
   = \frac{w_n}{w_c}.
 \end{equation}
Analogous to the crack band model presented by
Bazant~\cite{Bazant83}, Navas {\it et al.} \cite{Navas_2018_ES}
\cite{Navas_2017_ES} introduced a band width parameter $h_{\epsilon}$ in order to relate the crack opening with the strain.
Concerning this parameter, a typical value between two and
four times the maximum size of the aggregates is adopted in the case of
concrete as brittle material. The effective fracture strain
$\varepsilon_{\epsilon,f}$ is defined as the difference between the strain at crack initialization,
$\varepsilon_1(\vec{x}_p^{0})$, and the current strain, $\varepsilon_1(\vec{x}_p^{k+1})$, for a material point $p$. Also,
$\varepsilon_{\epsilon,f}$ can be represented as the current crack
opening $w_n$ within the band width, $h_{\epsilon}$. \DM{Check band width $h_{\epsilon}$ subindex or superindex? or both. See eq. 18,19 and 20} Therefore, 
\begin{equation}
  \label{eq:effective-fracture-strain}
  \varepsilon_{\epsilon,f} = \varepsilon_1(\vec{x}_p^{k+1}) -
  \varepsilon_1(\vec{x}_p^{0}) = \frac{w_n}{h_{\epsilon}}
\end{equation}
Introducing \eqref{eq:effective-fracture-strain} in
\eqref{eq:damaged-variable-chi}, the damage variable can be computed
as,
\begin{equation}
  \label{eq:damage-variable-chi-II}
\chi = \frac{\varepsilon_{\epsilon,f}\ h^{\epsilon}}{w_c}.  
\end{equation}
The function of $\chi$ presented in \eqref{eq:damage-variable-chi-II}
represents a linear softening behaviour. For a general case, the
damage variable can be expressed in terms of the following variables,
\begin{equation}
  \label{eq:damage-variable-chi-III}
  \chi = \chi(\varepsilon_{\epsilon,f}, h^{\epsilon}, f_t, w_c, G_f)
\end{equation}
Implementation details can be consulted in \ref{sec:eigens-algor-1}.

\subsection{$\epsilon$-neighborhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighborhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global definition of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. \DM{Intead, a} linked
list of tributary nodes is reconstructed with the closest node as
starting point at each time step \DM{deleted instead}. Therefore, each node is assigned to an
unique linked-list of those particles which are closest to it. This
approach allows to reconstruct easily a search list with the support
of the initial connectivity of the mesh and offers a suitable basis
for an easy implementation of other interpolation techniques with
minimum coding effort. A detailed explanation of the proposed
algorithm can be found in \ref{sec:node-linked-method-1}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cases of study and discussion}
\label{sec:3}

\DM{The proposed approach to overcome the two main shortcomings of EighenMPM \cite{Zhang_EE_2020} has been evaluated with "two benchmark tests"¿?. } \DM{DELETE: As noticed in \cite{Zhang_EE_2020}, the EigenMPM have two important
shortcomings}. The first one deals with the presence of stress instabilities provoked by the spatial discretization, appearing even
when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an
alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Section \ref{sec:3.3}, where experimental results
of a drop-weight impact test are compared with experimental ones as well as eigensoftening computations. \PNA{FALTA LO DE BUI}.

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}

The aim of the problem presented hereinafter is to assess the capability of the \acrshort{lme}
approximants to improve the result \textit{versus} the standard linear interpolation. The application consists of a square plate of size $H = 1$ containing
an initial edge crack of length $0.25 \cdot H$ loaded in a pure mode I
by displacement control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\PNA{UNIDADES}

\PNA{NOMBRAR ALGO A LOS CASOS ANALITICOS O DE MICHAEL ORTIZ}

The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the
Poisson's ratio \gls{nu} = 0.333, and critical energy-release rate
\gls{Gf} = 0.0001. This set of parameters has been previously employed by Pandolfi~\textit{et al.}~\cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretization are required. On the one hand,
a cartesian grid of nodes is considered with a nodal spacing value of
0.025. On the other hand, the plate will be modeled with a initial
layout of four particles per element and occupying the Gauss quadrature
positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows the
presence of wiggles in the reaction-displacements curve, in both loading and post-failure stages,
when linear interpolation technique is used. In contrast, \acrshort{lme} simulation does not
exhibit these spurious oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than
\acrshort{lme}. This can be attributed to imprecision in the stress
field due to grid-crossing phenomena.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
exhibits more clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. \PNA{Although, apparently, the differences in the obtained peaks are not extremely significant, it can be owing to the aforementioned hookean material employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive models will be employed, severe
%if hyperelastic materials are consider for future simulations,
inaccuracies may appear in the stress field that could affect dramatically to the
final result. Focusing on the post failure behavior, it can be seen how \acrshort{lme} produces soft stress
field evolution even after breaking.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: \PNA{once the energy is release dinamically by the crack, the waves propagate along the domain.}

\subsection{Drop-weight impact test}
\label{sec:3.2}
\PNA{This section is devoted to proof
the accuracy of the eigensoftening algorithm within the \acrshort{mpm} framework when the behavior
of quasi-brittle materials under dynamic loading is simulated. One of the most interesting examples of this
loading case is the three-point bending test on a concrete notched beam when it is conducted under impact loading. Indeed, this experimental test is appropriate to be reproduced numerically with the explicit solver proposed in this paper since
fracture occurs in a period of milliseconds and waves propagate fast along the high strength
materials involved in the aforementioned test.} Accordingly, small time steps are
required to achieve a numerically stable simulation.\\

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 881 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of discretization were required to assess
the objectiveness of the obtained results. The results here exposed
are from a discretization of 30616 material points and 6297 nodes. To
achieve better results, an unestructured grid layout has been adopted
focusing the minimum nodal spacing, 0.47 mm, in the middle of the beam and in
the impact surface of the hammer. For this simulation, \gls{gamma-lme} was
fixed to 6. Once the accuracy of the \acrshort{lme} approximants have been proven
in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam. %In the
%%\acrshort{mpm}, the implementation of this feature can be bypassed to obtain preliminary results. Nevertheless, this kind of algorithm could
%be implemented to get ever more accurate solutions. 
Despite the observed discrepancies, this is
out of the scope of this document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more realistic when modelled by a suitable contact algorithm instead of avoiding any vertical displacement}.\\

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:4}
\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrfull{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles provoked by inaccuracies in the stress integration. This spatial discretization was previously employed within the \acrshort{mpm} framework with excellent results under the elastic regime~\cite{Wobbes2020,Molinos2020}. The traditional time scheme of the \acrshort{mpm} has been improved with a Newmark Predictor Corrector scheme as well. In this research, the aforementioned techniques have helped to improve the
results in fracture propagation through the eigensoftening and eigenerosion algorithms. Further improvement may be reached by adapting \acrshort{lme} through the
deformation gradient\MMP{CITA AQUI LOS QUE LO HACEN}. \textit{A priori} this would enhance the localisation properties of the proposed algorithms in more sophisticated domains.

Finally, for better capturing the measurement of forces made in the laboratory, more realistic contact tools have to be developed. Navas~\textif{et al.} demonstrate the importance of the stiffness of the support in order to obtain accurate results of the reaction forces. Interesting contact algorithms within the \acrshort{mpm} methodology are available in the literature \MMP{CITA ALGUNO Y SI SON CONSERVATIVOS EN ENERGIAS, MEJOR}. }

\section*{Acknowledgements}
The financial support to develop this research from the Ministerio de
Ciencia e Innovaci\'on, under Grant No. BIA-2016-76253 is greatly
appreciated. The first and the second authors also acknowledge the
fellowship Fundaci\'on Agust\'in de Betancourt and Juan de la Cierva
(FJCI-2017–31544) respectively.

\appendix
%\addappheadtotoc
%\appendixpage
%\renewcommand{\theequation}{\Alph{section}.\arabic{equation}}

\section{Explicit Newmark Predictor-Corrector algorithm}
\label{sec:expl-pred-corr}
\begin{algorithm}[H]
  \DontPrintSemicolon
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update mass matrix :} $ \tens{m}_{I} = N_{Ip}^{k}\ m_p$ \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Predictor :}
    \begin{equation*}
      \vec{v}_I^{pred} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Impose essential boundary conditions :} At the fixed
    boundary, set $\vec{v}_{I}^{pred} = 0$.\; 
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Deformation tensor increment calculation :}
    \begin{equation*}
      \Delta \tens{\varepsilon}_{p}^{k+1} = \Delta t\
        \dot{\tens{\varepsilon}_{p}}^{k+1} = \Delta t\ \left[ \vec{v}_{I}^{pred} \otimes
        \Grad{N_{Ip}^{k+1}} \right]^s
    \end{equation*} \;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Update the density field :} $\rho_p^{k+1} =
    \frac{\rho_p^k}{1 + \mathit{tra}\left[\Delta\tens{\varepsilon}_{p}^{k+1}\right]}.$\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Compute stress field and update damage parameter}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Balance of forces calculation :} Calculate the total grid
    nodal force $\vec{f}_{I}^{k+1} = \vec{f}_{I}^{int,k+1} + \vec{f}_{I}^{ext,k+1}$.\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Explicit Newmark Corrector :}
    \begin{equation*}
      \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}  
    \end{equation*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Update particles lagrangian quantities :}
    \begin{align*}
      &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}
    \end{align*}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \textbf{Reset nodal values}\;
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
    \label{alg-epc}
    \caption{Explicit Newmark Predictor-Corrector scheme}
\end{algorithm}

\section{Eigensoftening Algorithm}
\label{sec:eigens-algor-1}

\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{For each $p$, $\epsilon$-neighbourhood, $f_{t,p}$, $h_{\epsilon,p}$, $w_c$}
  \KwOutput{Return damage parameter $\chi := \{\chi_p\}$}
  % \KwData{Testing set $x$}
  $\chi_p \leftarrow \chi_p^{k}$ \;
  \For{$p$ to $N_p$}
  {
    \If{$\chi_p = 0$ \AND $\epsilon_{f,p} = 0$}
    {
      \For{$q \in B_{\epsilon,p}$}
      {
        $\sigma_{q,I} \leftarrow \text{getEigenvaluesOf}(\sigma_{q})$ \;
        \If{$\chi_q < 1$}
        {
          $\sum m_p\sigma_{p,I} \leftarrow \sum m_p\sigma_{p,I} + m_q\sigma_{q,I}$ \;
        }
        $m_p \leftarrow m_p + m_q$ \;
      }
      $\sigma_{p,\epsilon} \leftarrow \frac{1}{m_p} \sum m_p\sigma_{p,I}$\;
      \If{$\sigma_{p,\epsilon} > f_{t,p}$}
      {
        $\epsilon_{f,p} = \epsilon_{I,p}$ \;  
      }
      \ElseIf{$\chi_p \neq 1$ \AND $\epsilon_{f,p} > 0$}
      {
        $\chi_p^{k+1} \leftarrow \min\Big \{1 , \max \{\chi_p^{k},
        \frac{(\epsilon_{I,p}- \epsilon_{f,p})\ h_{\epsilon,p}}{w_c}
        \} \Big \}$ \;
      }
    }
  }
  \caption{Compute damage parameter $\chi_p^{k+1}$}
  \label{alg-eigens}
\end{algorithm}

\section{Node-linked method for $\epsilon$-neighborhood reconstruction}
\label{sec:node-linked-method-1}
\begin{algorithm}[H]
  \DontPrintSemicolon
  \KwInput{$I_0$ : Closest node for each particle}
  \KwInput{$I_{Nodes}$ : List of nodes close to each node}
  \KwInput{$Particles_I$ : List of particles close to each node}
  \KwOutput{$\text{neighborhood}_{\epsilon}$ : $\epsilon$-neighborhood for each particle}
  \For{$p$ to $N_p$}
  {
    For each $p$ get $I_0$\;
    Eval $I_{Nodes}(I_0)$ \;
    \For{$k \in I_{Nodes}(I_0)$}
    {
      $\text{neighborhood}_{\epsilon}^{k+1}$ = $\text{neighborhood}_{\epsilon}^{k} \cup Particles_I(k)$
    }
    \For{$p_{\epsilon} \in \text{neighborhood}_{\epsilon}$}
    {
      \If{$distance(p_{\epsilon},p) > \epsilon $}
      {
        pop $p_{\epsilon}$ from neighborhood$_{\epsilon}$\;
      }      
    }

  }
  \label{alg-node-link}
  \caption{$\epsilon$-neighborhood reconstruction algorithm}
\end{algorithm}

% name your BibTeX data base
\bibliography{Biblio}

\end{document}