### 2020-06-30 Cover-letter.tex

\batchmode
\documentclass[10pt]{letter}

\usepackage{mathptm}
\usepackage{fullpage}

\pagestyle{empty}
\begin{document}

Mr. Miguel Molinos\\
ETSI Caminos, Canales y Puertos\\
Technical University of Madrid\\
C/ Prof. Aranguren 3\\
28040 Madrid, Spain\\
Tel. N.:+34 910674396\\
email address: m.molinos@alumnos.upm.es\\

\vspace{.2in}

30 June 2020

\vspace{.3in}

Editor\\
Computers and Structures\\

Dear Sir:

\vspace{.2in}

Please find enclosed the electronic submission of the
\textbf{research article} entitled \textit{``LME-MPM applied to quasi-brittle fracture''} by Miguel Molinos, Pedro Navas, Diego Manzanal and Manuel Pastor. We would like to have this manuscript reviewed by \textbf{Computers and Structures}.

The authors of this paper warrant that this manuscript represents
original work that is not being considered for publication, in
whole or in part, in another journal, book, conference proceedings
or other publications. No already published work can be found in
this document, and all used sources have been cited in the
\textit{References} section. In addition, this manuscript is one
of a kind. All the authors have contributed substantially to the
manuscript and approved the final submission.

Should you need to contact us, please, use the details given above.

\vspace{.1in}

Sincerely,

\vspace{.3in}

Miguel Molinos

\end{document}