### 2020-07-30 Author-Novelty.tex
Cambio de revista

\documentclass[12pt,a4paper]{article}
\usepackage{mathptm}

\pagenumbering{gobble}

\title{LME-MPM applied to quasi-brittle fracture}
\author{Miguel Molinos, Pedro Navas, Diego Manzanal and Manuel Pastor}
%\date{}                                           % Activate to display a given date or no date
\begin{document}
\maketitle

\centering
\section*{Highlights}
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\begin{itemize}
\item Fracture mechanics within the MPM is a challenging research field

\item The Eigensoftening algorithm is an improvement to simulate quasi-brittle materials

\item The LME shape functions together with a NPC time integration enhance the standard MPM

\item The Eigensoftening-MPM has been validated against analytical and experimental results

\end{itemize}

\end{document}